---
title: "TMWP_MatrixMath_Experimentation.Rmd"
author: "Created by Mitch, github ID: TheMitchWorksPro"
date: "March 20, 2017"
output: html_document
---

```{r firstCell, eval=FALSE, echo=FALSE}

# files to load (when created):  
#   to reload environment with variables and previously run output:
#   load("tbd.RData")  # recreates the working environment (assuming we saved it)

```

### TOC

In this document:

  - [Outer Product Experiment - Numeric Vector](#problem1)
  - [Outer Product Experiment - Character Vector](#problem1b)
  - [Inner Product Experiment - Numeric Vector](#problem2)

<a id="problem1" name="problem1"></a>

### Outer Product Experiment One - Numeric Vectors

These cells test some [outer product code from a post on nabble.com](http://r.789695.n4.nabble.com/Outer-product-from-matrix-by-row-and-a-vector-td4639525.html). Then we conduct some experiments with it.  [This link](https://stat.ethz.ch/R-manual/R-devel/library/base/html/matmult.html) was also helpful for code that comes later in this document.

```{r problem1StartsHere, echo=TRUE, eval=TRUE}

# http://r.789695.n4.nabble.com/Outer-product-from-matrix-by-row-and-a-vector-td4639525.html

# code from stack overflow ... only varaible names have been altered
# this is just a test to understand the code

aVec <- matrix(c(1:6),ncol=3)
aVec
bVec <- c(1,2,3,4)
bVec 
out <- aVec[, 1] + aVec[, 3] %o% (1/bVec)
out

result <- matrix(rep(0, times=dim(aVec)[1]*length(bVec)),nrow=dim(aVec)[1]) 
  # this buildes a vector full of zeroes probably of desired dimensions 
  # for next steps ...

result

```

```{r soexamplecont2}
for(ii in 1:dim(aVec)[1]){ 
  for(jj in 1:length(bVec)){ 
    result[ii,jj] <- aVec[ii,1]+aVec[ii,3]/bVec[jj] 
  } 
} 

max(abs(out - result))

```

```{r soexamplecont4}
identical( t( apply(aVec, 1, function(x) { x[1] + x[3]/ bVec } ) ), result) 

```

Now we try to build our own.  This code requires a numeric vector as input, so this version explores the outer product code using numeric vectors:

```{r myCell1_pasteNums, echo=TRUE, eval=TRUE}

paste_nums <- function(number_vec) {
  # do something here
  return(t(as.matrix(number_vec)) %o% as.matrix(number_vec)) 
}

myVec1 <- c(1, 2, 3)
myVec1

```

```{r pastenums2}
outObj <- paste_nums(myVec1)
outObj

```

To get the sum of the outer product, we could just do this:
```{r pastenums3}
sum(paste_nums(myVec1))

```

<a id="problem1b" name="problem1b"></a>

### Outer Product Experiment Two - Character Vector

but how do we do it to a character vec like: 'a', 'b', 'c'? The above code works with numbers only.
Desired output is like "outObj" but on our character vector.  First a quick experiment

```{r textExp1}

myVec2 <- c('a', 'b', 'c')
myVec2

paste("text", myVec2)

```

this solution uses an [sappy](http://stackoverflow.com/questions/3505701/r-grouping-functions-sapply-vs-lapply-vs-apply-vs-tapply-vs-by-vs-aggrega) function mapping to achieve the same effect with a text vector:
```{r charOutsolution1}

paste_chars <- function(char_vec) {
    return(sapply(char_vec, paste, arg1=char_vec, sep=""))
}

paste_chars(myVec2)

```

more testing of our solution ... outer product of a character vector w/ itself ...

```{r charVecSolutionCell2}

myVec3 <- c('x', 'y', 'z', 'a', 'n', 'g', 'word')
myVec3

paste_chars(myVec3)

```

<a id="problem2" name="problem2"></a>

### Inner Product - Numeric


`%i%` is not valid and `%in%` means find the values of one vector within another returning a boolean vector of which values were found.  So how do we do inner product?

```{r inTest1}
aVec2 <- matrix(c(1:6),ncol=3)
aVec2
bVec2 <- matrix(c(6:11), ncol=2)
bVec2

library(grDevices)
inProd <- aVec2 %*% bVec2
inProd

```

A quick explanation of matrix math performed above:

 (1\*6) + (3\*7) + (5\*8) = *67* &nbsp;&nbsp;&nbsp;&nbsp; (1\*9) + (3\*10) + (5\*11) =  *94*  
 (2\*6) + (4\*7) + (6\*8) = *88* &nbsp;&nbsp;&nbsp;&nbsp; (2\*9) + (4\*10) + (6\*11) = *124*  

```{r finalAnswerInnerProd}
# inner product
sum(inProd)

```

